Efficient pedigree recording for fast population genetics simulation

### 2017-12-30 forwards_paper.tex
Add simupop reqs

    \item
        In Python, using the \simupop{} library interfacing with a custom Python library (\ftprime{}),
        using a discrete but large number of sites.
        % $\mu$ and $r$ define per-base-pair mutation rate and we
        % defined the length of the simulated sequence, in base pairs, as $L = \tfrac{\rho}{4N \mu}$.
        % We varied $N \in (1000, 10000)$, $\rho \in (10^3, 10^4)$.

@@

\begin{figure}
    \begin{center}
        \includegraphics{simplify-state-diagram}
    \end{center}
    \caption{
        A depiction of the state of the simplification algorithm
        at each point in time,
        in the example of Figure~\ref{fig:method_diagram}.
        \label{fig:simplify_state}
    }
\end{figure}

@@

The Tables API is primarily designed to facilitate efficient interchange of
data between programs or between different modules of the same program.
% we are capitalizing Tables API but not the more generic "tables".
% \jda{should capitalize Tables, or not, as above}
Following current best practises
\plr{citations: Apache arrow, etc},
data is stored in a columnar format,
which allows for very efficient interchange of large amounts of numerical data.
In principle, this enables zero-copy semantics,
where a data consumer can read the information directly
from the memory of a producer without incurring the overhead of a copy.
Our implementation uses the numpy C API \citep{numpy_C_api} to efficiently copy
data from Python into the low-level C library used to manipulate tree sequences.
Since tables are encoded as contiguous arrays of numeric data,
this achieves data transfer rates that would be impossible under
any text-based approach while retaining excellent portability.

This also provides basic input and output operations via numpy's
array interface, which provides a great deal of flexibility and efficiency,
and makes it straightforward to transfer data from sources
such as HDF5 \citep{hdf5}, Dask, or Zarr.
(For small scale data and debugging purposes, a simple text format is also supported.)
% PLR: this was mentioned above, and belongs in the documentation, I think.
% The API also provides functions to sort tables
% to ensure that the records are in the form required to input
% into an \msprime{} tree sequence object. (Simplification may also be required.)

@@

>>>>

    \item
        In Python, using the \simupop{} library interfacing with Python code
	(via our package \ftprime{}),
        using a discrete but large number of sites.
        % $\mu$ and $r$ define per-base-pair mutation rate and we
        % defined the length of the simulated sequence, in base pairs, as $L = \tfrac{\rho}{4N \mu}$.
        % We varied $N \in (1000, 10000)$, $\rho \in (10^3, 10^4)$.

@@

\begin{figure}
    \begin{center}
        \includegraphics{simplify-state-diagram}
    \end{center}
    \caption{
        A depiction of the state of the simplification algorithm
        at each point in time,
        in the example of Figure~\ref{fig:method_diagram}A.
        As in that figure, blue numbers denote the crossing over location in a meiosis --
        for instance, $G$ and $H$ were parents to $I$,
        who inherited the left 70\% of the chromosome from $G$ and the remainder from $H$.
        \label{fig:simplify_state}
    }
\end{figure}

@@

The Tables API is primarily designed to facilitate efficient interchange of
data between programs or between different modules of the same program.
% we are capitalizing Tables API but not the more generic "tables".
% \jda{should capitalize Tables, or not, as above}
Following current best practises
\plr{citations: Apache arrow, etc},
data is stored in a columnar format,
which allows for very efficient interchange of large amounts of numerical data.
In principle, this enables zero-copy semantics,
where a data consumer can read the information directly
from the memory of a producer without incurring the overhead of a copy.
\jda{TODO citation missing}
Our implementation uses the numpy C API \citep{numpy_C_api} to efficiently copy
data from Python into the low-level C library used to manipulate tree sequences.
Since tables are encoded as contiguous arrays of numeric data,
this achieves data transfer rates that would be impossible under
any text-based approach while retaining excellent portability.

This also provides basic input and output operations via numpy's
array interface, which provides a great deal of flexibility and efficiency,
and makes it straightforward to transfer data from sources
such as HDF5 \citep{hdf5}, Dask, or Zarr.
\jda{TODO citation missing}
(For small scale data and debugging purposes, a simple text format is also supported.)
% PLR: this was mentioned above, and belongs in the documentation, I think.
% The API also provides functions to sort tables
% to ensure that the records are in the form required to input
% into an \msprime{} tree sequence object. (Simplification may also be required.)

@@

\subsection{Requirements for the \simupop{} implementation}

The simulations are implemented the Python package \ftprime{} 0.0.6rc0
which is available at \url{https://www.github.com/ashander/ftprime}.
Its requirements are encoded in an \texttt{environment.yml} file, compatible
with conda, but \msprime{} must be updated to the version listed below via pip
from GitHub \jda{TODO add command}.
In brief the requirements are:

\begin{enumerate}
    \item \msprime{} \jda{TODO add version from KRT's dev machine}
    \item \simupop{} 1.1.8.3
    \item python 3.5
\end{enumerate}
